---
output: html_document
---

## Contact Tracing for COVID-19

Updated `r Sys.Date()`

```{r setup, include=FALSE}
  knitr::opts_chunk$set(echo = FALSE, fig.align="center")
  library(tidyverse)
  library(gridExtra)

  source("../R/ContactTracing.R")

```


### Data Sources

#### Natural history

[Incubation period:](https://annals.org/aim/fullarticle/2762808/incubation-period-coronavirus-disease-2019-covid-19-from-publicly-reported) 5.1d (log-mean 1.621, log-SD 0.418).

[Distribution of Infectiousness](https://www.nature.com/articles/s41591-020-0869-5/figures/1): relative infectiousness from time of symptom onset follows a gamma distribution with shape 2.12, rate 0.69, amd offest 2.31.

This is consistent with other analyses that find a median serial interval of [4.0 days](https://wwwnc.cdc.gov/eid/article/26/6/20-0357_article), [4.6 days](https://www.sciencedirect.com/science/article/pii/S1201971220301193), [5.1 days](https://www.thelancet.com/journals/laninf/article/PIIS1473-3099(20)30230-9/fulltext),  

#### Contact Data

[The POLYMOD study](https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.0050074) estimates an average of 13.4 contacts per participant per day, including children, though there is variance across countries (e.g., mean 8 in Germany, mean 19.8 in Italy). There was some skew to the data, which also varied by country. Between 20 - 30% of contacts were made in the household.

A [survey in the UK](https://www.medrxiv.org/content/10.1101/2020.03.31.20049023v1?%253fcollection=) found that, from 24 - 26 March 2020, adults reported 3 (IQR 1, 4) physical and non-physical contacts, and that 57.6% of contacts occurred at home.

A [survey in California, New York, Arizona, and Massachusetts, USA](https://www.medrxiv.org/content/10.1101/2020.04.13.20064014v1) found that, from 22 March - 8 April 2020, adults reported an average of 2.7 contacts (median 2) conversational contacts, using the same definition as the POLYMOD study. There were few contacts (mean ~1) outside of the household; 50% of respondents had no contacts outside of the household, and the vast majority that did have contact outside the household reported only 1 or 2 contacts.

Preliminary data from contact tracing efforts in Maryland during the stay-at-home order have indicated that roughly half of adults report no contacts, and those that do report contacts almost always report fewer than 4 contacts. Up to 25% of cases cannot be reached once test results are available.

### Scenarios

We list a table below of the key scenarios modeled below.

|    | Scenario                                     | Description                                                  |
|---:|:--------------------------------------------:|:------------------------------------------------------------:|
| 1  | Traditional, test-based contact isolation    | Passive surveillance isolation on average 4.6d after symptom onset; contact tracing isolation on average 2.7d after symptom onset |
| 2  | Rapid, test-based contact isolation          | Scenario 1, but individuals are isolated on average 2 days faster |
| 3  | Household quarantine upon index confirmation | Scenario 1, but all household contacts are quarntined on day of test confirmation/isolation of index case |
| 4  | Household quarantine upon index detection    | Index cases and household contacts are quarantined on day of index case symptom onset |


### Key Questions

1. **How effective is 'traditional' contact tracing?**

Here, we explore a scenario where 10% of infectious individuals are identified through passive surveillance and are contacted to begin isolation and contact tracing. We assume that moderate social distancing is in place, such that each individual would report on average 5 contacts, with average 60% of contacts occurring within the home. On average, cases detected through symptom-based passive surveillance are isolated within 4.6 days of symptom onset. Cases identified through contact tracing on average are isolated within 2.7 days. We estimate that 75% of household contacts of infected cases can be reached and successfully isolated and that 50% of non-household contacts can be reached and successfully traced. We assume that household contacts are 4 times more likely to be infected than non-household contacts.

```{r}

nsims = 10000
I0 = 1000
p_scn1 <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(0.77, 0.67),
               delay_parC = c(0.77, 0.67),
               ind_onset_H = TRUE,
               ind_onset_C = TRUE)

scn1 <- run_tracing_sim(nsims, I0, p_scn1)
calcRe_sim(gen1 = scn1$I1_total, gen2 = scn1$I2_total)
calcRe_sim(gen1 = scn1$I1_passive, gen2 = scn1$I2_passive)
calcRe_exact(p_scn1)

```

2. **What happens if we shorten the time it takes to identify cases and trace contacts?**

This scenario is the same as Scenario 1 described above, but we assume that we are able to isolate cases and their contacts on average 2 days earlier (on average, 2.6 days and 0.7 days since symptom onset, respectively). 

```{r}
nsims = 10000
I0 = 1000
p_scn2 <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(0.65, 0.78),
               delay_parH = c(-0.58, 0.67),
               delay_parC = c(-0.58, 0.67),
               ind_onset_H = TRUE,
               ind_onset_C = TRUE)

scn2 <- run_tracing_sim(nsims, I0, p_scn2)
calcRe_sim(gen1 = scn2$I1_total, gen2 = scn2$I2_total)
calcRe_sim(gen1 = scn2$I1_passive, gen2 = scn2$I2_passive)
calcRe_exact(p_scn2)

```

3. **What happens if we quarantine household members at the same time as we isolate index cases after positive test?**

This scenario is the same as Scenario 1 described above, but we now assume that 75% of household members are successfully quarantined on the same day the index case is isolated (4.6 days after index case symptom onset or about 1 day *before* symptom onset of infected household contacts, on average). Non-household contacts are still manually traced following test confirmation, and on average are isolated 2.7 days after their symptom onset.

```{r}

nsims = 10000
I0 = 1000
p_scn3 <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.2,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(1.12, 0.67), # indexed from infection
               delay_parC = c(0.77, 0.67), # indexed from onset
               ind_onset_H = FALSE,
               ind_onset_C = TRUE)

scn3 <- run_tracing_sim(nsims, I0, p_scn3)
calcRe_sim(gen1 = scn3$I1_total, gen2 = scn3$I2_total)
calcRe_sim(gen1 = scn3$I1_passive, gen2 = scn3$I2_passive)
calcRe_exact(p_scn3) 

```

4. **What happens if individuals and their household are asked to isolate when they first contact a healthcare provider, prior to testing or confirmation?**

This scenario is the same as Scenario 3 described above, but we now assume that infectious individuals are isolated, on average, on the day of symptom onset, when they might first contact a healthcare provider. Household contacts are quarantined on the same day which, following the serial interval of SARS-CoV-2, will be on average about 5 days *before* symptom onset of infected household contacts. Non-household contacts are still manually traced following test confirmation, and on average are isolated 2.7 days after their symptom onset.

```{r}

nsims = 10000
I0 = 1000
p_scn4 <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.2,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(-3, 1),
               delay_parH = c(-3, 0.5),
               delay_parC = c(0.77, 0.67),
               ind_onset_H = FALSE,
               ind_onset_C = TRUE)

scn4 <- run_tracing_sim(nsims, I0, p_scn4)
calcRe_sim(gen1 = scn4$I1_total, gen2 = scn4$I2_total)
calcRe_sim(gen1 = scn4$I1_passive, gen2 = scn4$I2_passive)
calcRe_exact(p_scn4) 

```

5. **How quickly do contacts need to be identified, before there are few additional benefits to quarantine?**

Below, we plot $R_{eff, p}$, the effective reproductive number among contacts of cases detected through passive surveillance, and $R_{eff}$, the effective reproductive number considering all infected individuals, while varying the median time from infection to quarantine for household and non-household contacts of a confirmed case. We assume (as in Scenario 4) that cases are, on average, passively detected on day of symptom onset and that 75% of household contacts and 50% of non-household contacts are successfully quarantined.

```{r gammaC gammaH, fig.width=8}
p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.2,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 1,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(-3, 1),
               delay_parH = c(0, 0.5),
               delay_parC = c(0, 0.5),
               ind_onset_H = FALSE,
               ind_onset_C = FALSE)

delays <- c(0.05, seq(0.1, 10, 0.1))
out <- matrix(NA, nrow=length(delays),ncol=5)
out[,1] <- delays
for(i in 1:length(delays)){
  p_base$delay_parH[1] <- p_base$delay_parC[1] <- log(delays[i])
  out[i, 2:5] <- calcRe_exact(p_base)
}
colnames(out) <- c("delay", names(calcRe_exact(p_base)))

par(mfrow=c(1,2))
plot(out[,'delay'], out[,'Re_I1_passive'],
     xlab="Time from contact infection to quarantine",
     ylab="R_eff,p",
     pch=16,
     ylim=c(0.75, 2.5))
abline(h=1, lty=2)


plot(out[,'delay'], out[,'Re_I1'],
     xlab="Time from contact infection to quarantine",
     ylab="R_eff",
     pch=16,
     ylim=c(0.75, 2.5))
abline(h=1, lty=2)

```


6. **What proportion of contacts need to be identified to achieve a certain reduction in transmission?**


```{r propC, eval=FALSE, include=FALSE}

# Below, we plot $R_{eff, p}$, the effective reproductive number among contacts of cases detected through passive surveillance, and $R_{eff}$, the effective reproductive number considering all infected individuals, while varying the proportion of non-houshold contacts that are isolated in each of the four scenarios described above.

p_list = list(p_scn1, p_scn2, p_scn3, p_scn4)
propC <- seq(0, 1, 0.025)
outC <- matrix(NA,ncol=6)
for(i in 1:length(propC)){
  for(j in 1:length(p_list)){
    p_list[[j]]$rhoC <- propC[i]
    outC <- rbind(outC,
                  c(propC[i], j, calcRe_exact(p_list[[j]])))
  }
}
outC <- data.frame(outC[-1,])
colnames(outC) <- c("propC", "scn", names(calcRe_exact(p_base)))

p1 <- ggplot(outC, aes(x=propC, y=Re_I1_passive, group=as.factor(scn), color=as.factor(scn))) +
        geom_line() +
        geom_hline(aes(yintercept=1), lty=2, col='gray') +
        theme_minimal() +
        xlab("proportion non-household contacts isolated") +
        ylab("R_eff, p") +
        scale_color_discrete(name="Scenario")
  
p2 <- ggplot(outC, aes(x=propC, y=Re_I1, group=as.factor(scn), color=as.factor(scn))) +
        geom_line() +
        geom_hline(aes(yintercept=1), lty=2, col='gray') +
        theme_minimal() +
        xlab("proportion non-household contacts isolated") +
        ylab("R_eff") +
        scale_color_discrete(name="Scenario")

grid.arrange(p1, p2, nrow=1)

```

This plot shows $R_{eff}$, the effective reproductive number considering all infected individuals, while varying the proportion of transmission due to cases identified through passive surveillance (including asymptomatic or non-clinically apparent cases). We have also included a horizontal line at $R_eff$ = 1.25, a possible target if we assume a further 20% reduction in transmission through social distancing.

```{r prop1}

p_list = list(p_scn1, p_scn2, p_scn3, p_scn4)
propP <- seq(0, 1, 0.025)
outP <- matrix(NA,ncol=6)
for(i in 1:length(propP)){
  for(j in 1:length(p_list)){
    p_list[[j]]$rho1 <- propP[i]
    outP <- rbind(outP,
                  c(p_list[[j]]$rho1, j, calcRe_exact(p_list[[j]])))
  }
}
outP <- data.frame(outP[-1,])
colnames(outP) <- c("propP", "scn", names(calcRe_exact(p_base)))

ggplot(outP, aes(x=propP, y=Re_I1, group=as.factor(scn), color=as.factor(scn))) +
        geom_line() +
        geom_hline(aes(yintercept=1), lty=2, col='gray') +
        geom_hline(aes(yintercept=1.25), lty=2, col='darkgray') +
        theme_minimal() +
        xlab("prop. transmission from cases detected in passive surveillance") +
        ylab("R_eff") +
        scale_color_discrete(name="Scenario")

```


7. **How does the importance of tracing household vs. non-household contacts change with the relative infectiousness of household vs non-houshold contacts?**

There is evidence that attack rates among household contacts are 6 times higher than attack rates among non-household close contacts ([Bi et al.](https://www.thelancet.com/journals/laninf/article/PIIS1473-3099(20)30287-5/fulltext#tbl3)). Below, we plot $R_{eff,p}$ with different values of $\nu$ = 1, 4, 6, 10, the relative reproductive number in household contacts compared to non-household contacts, across varying levels of $\rho_H$, the proportion of household contacts successfully quarantined, and $\rho_C$, the proportion of non-household contacts successfully quarantined. We show results for Scenario 2 (rapid test-based isolation and contact symptom-based quarantine) and Scenario 4 (immediate symptom-based isolation and household quarantine). The red line shows $R_{eff,p}=1$

```{r nu comparison, fig.width = 8, fig.height = 13}

p_list = list(p_scn2, p_scn4)
propC <- seq(0, 1, 0.01)
propH <- seq(0, 1, 0.01)
relInfs <- c(1, 4, 6, 10)
out <- matrix(NA,ncol=8)
for(k in 1:length(p_list)){
  for(i in 1:length(propC)){
    p_list[[k]]$rhoC <- propC[i]
    for(j in 1:length(propH)){
    p_list[[k]]$rhoH <- propH[j]
    for(n in 1:length(relInfs)){
        p_list[[k]]$nu <- relInfs[n]
        out <- rbind(out,
                    c(p_list[[k]]$rhoC, p_list[[k]]$rhoH, p_list[[k]]$nu, k, calcRe_exact(p_list[[k]])))
      }
    }
  }
}

out <- data.frame(out[-1,])
colnames(out) <- c("propC", "propH", "nu", "scn", names(calcRe_exact(p_base)))
out$scn_label[out$scn==1] <- "Scn 2"
out$scn_label[out$scn==2] <- "Scn 4"

ggplot(out, aes(x=propC, y=propH, fill=Re_I1_passive)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1_passive, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  facet_grid(nu~scn_label) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff,p") +
  xlab("Prop. non-household contacts quarantined") +
  ylab("Prop. household contacts quarantined")

```

8. **How does the effective reproductive number change as a function of the delays to quarantine of household and non-household contacts?**

Below, we plot $R_{eff,p}$ across different average delays from exposure to quarantine of household and non-household contacts of confirmed cases underotherwise Traditional contact tracing (Scenario 1). This represents an optimistic scenario where 90% of household contacts and 75% of non-household contacts are traced and isolated. 

```{r scn1 gamma comparison, fig.width = 5, fig.height = 13}

p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0.9,
               rhoC = 0.75,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(-3, 0.67),
               delay_parC = c(-3, 0.67),
               ind_onset_H = FALSE,
               ind_onset_C = FALSE)

delaysC <- seq(0.075, 10, 0.075)
delaysH <- seq(0.075, 10, 0.075)
relInfs <- c(1, 4, 6, 10)
out <- matrix(NA,ncol=7)
for(i in 1:length(delaysC)){
    p_base$delay_parC[1] <- log(delaysC[i])
  for(j in 1:length(delaysH)){
    p_base$delay_parH[1] <- log(delaysH[j])
    for(n in 1:length(relInfs)){
        p_base$nu <- relInfs[n]
        out <- rbind(out,
                    c(delaysC[i], delaysH[j], p_base$nu, calcRe_exact(p_base)))
    }
  }
}


out <- data.frame(out[-1,])
colnames(out) <- c("delayC", "delayH", "nu", names(calcRe_exact(p_base)))

ggplot(out, aes(x=delayC, y=delayH, fill=Re_I1_passive)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1_passive, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  facet_grid(nu~.) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff,p") +
  xlab("Time from non-HH contact exposure to quarantine") +
  ylab("Time from HH contact exposure to quarantine")

```

The following figure (left) shows the trade-offs between proportion of household contacts quarantined and the average delay from infection to isolation of infected household contacts. This assumes an optimistic scenario where 75% of non-household contacts are quarantined on the same day that the index case is isolated (4.6 days after index case symptom onset).

The figure on the right shows the equivalent plot, focusing on non-household contacts, for the trade-offs between proportion quarantined and average delays from infection to isolation of non-household contacts. This assumes an optimistic scenario where 90% of household contacts are quarantined on the same day that the index case is isolated (4.6 days after index case symptom onset).


```{r scn1 rhoH, fig.width = 5, fig.height = 10}

p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0,
               rhoC = 0.75,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(-3, 0.67),
               delay_parC = c(1.12, 0.67),
               ind_onset_H = FALSE,
               ind_onset_C = FALSE)

propH <- seq(0, 1, 0.025)
delaysH <- seq(0.075, 10, 0.075)
relInfs <- c(1, 4, 6, 10)
out <- matrix(NA,ncol=7)
for(i in 1:length(delaysH)){
    p_base$delay_parH[1] <- log(delaysH[i])
  for(j in 1:length(propH)){
    p_base$rhoH <- propH[j]
    for(n in 1:length(relInfs)){
        p_base$nu <- relInfs[n]
        out <- rbind(out,
                    c(delaysH[i], p_base$rhoH, p_base$nu, calcRe_exact(p_base)))
    }
  }
}


out <- data.frame(out[-1,])
colnames(out) <- c("delayH", "rhoH", "nu", names(calcRe_exact(p_base)))

pH <- ggplot(out, aes(x=delayH, y=rhoH, fill=Re_I1_passive)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1_passive, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  facet_grid(nu~.) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff,p") +
  xlab("Time from HH contact exposure to quarantine") +
  ylab("Proportion HH contacts quarantined")

```


```{r scn4 gamma comparison, fig.width = 12, fig.height = 13}

p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0,
               rhoC = 0.75,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(1.12, 0.67),
               delay_parC = c(-3, 1),
               ind_onset_H = FALSE,
               ind_onset_C = FALSE)

delaysC <- seq(0.075, 10, 0.075)
propC <- seq(0, 1, 0.025)
relInfs <- c(1, 4, 6, 10)
out <- matrix(NA,ncol=7)
for(i in 1:length(delaysC)){
    p_base$delay_parC[1] <- log(delaysC[i])
  for(j in 1:length(propC)){
    p_base$rhoC <- propC[j]
    for(n in 1:length(relInfs)){
        p_base$nu <- relInfs[n]
        out <- rbind(out,
                    c(delaysC[i], p_base$rhoC, p_base$nu, calcRe_exact(p_base)))
    }
  }
}


out <- data.frame(out[-1,])
colnames(out) <- c("delayC", "rhoC", "nu", names(calcRe_exact(p_base)))

pC <- ggplot(out, aes(x=delayC, y=rhoC, fill=Re_I1_passive)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1_passive, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  facet_grid(nu~.) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff,p") +
  xlab("Time from non-HH contact exposure to quarantine") +
  ylab("Proportion non-HH contacts quarantined")

grid.arrange(pH, pC, nrow=1)
```


```{r propP comparison, eval=FALSE, include=FALSE}

# Here, we plot $R_{eff}$ across different proportion of cases detected through passive surveillance and average delays from symptom onset to isolation under a traditional strategy in which 90% of household contacts and 75% of non-household contacts are quarantined on average within 2.7 days of symptom onset.

p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0.75,
               rhoC = 0.5,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(1.22, 0.78),
               delay_parH = c(0.77, 0.67),
               delay_parC = c(0.77, 0.67),
               ind_onset = TRUE)

delaysP <- seq(0.05, 4, 0.05)
propP <- seq(0, 1, 0.025)
out <- matrix(NA,ncol=6)
for(i in 1:length(delaysP)){
    p_base$delay_parP[1] <- log(delaysP[i])
  for(j in 1:length(propP)){
    p_base$rho1 <- propP[j]
    out <- rbind(out,
                c(delaysP[i], p_base$rho1, calcRe_exact(p_base)))
  }
}


out <- data.frame(out[-1,])
colnames(out) <- c("delayP", "rhoP", names(calcRe_exact(p_base)))

ggplot(out, aes(x=delayP, y=rhoP, fill=Re_I1)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff") +
  xlab("Time from symptom onset to index isolation") +
  ylab("Proportion index cases detected")

```


```{r scn4 propP, eval=FALSE, include=FALSE}

p_base <- list(R0 = 2.5,
               theta = 0.1,
               N = 5,
               thetaN = 0.1,
               alphaH = 0.60,
               rho1 = 0.1,
               rhoH = 0.90,
               rhoC = 0.75,
               nu = 4,
               inf_par = c(2.1157790, 0.6898583, 2.3066912),
               inc_par = c(1.621, 0.418),
               delay_parP = c(-3, 1),
               delay_parH = c(-3, 1),
               delay_parC = c(0.77, 0.67),
               ind_onset = FALSE)


delaysP <- seq(0.05, 4, 0.05)
propP <- seq(0, 1, 0.025)
out <- matrix(NA,ncol=6)
for(i in 1:length(delaysP)){
    p_base$delay_parP[1] <- log(delaysP[i])
  for(j in 1:length(propP)){
    p_base$rho1 <- propP[j]
    out <- rbind(out,
                c(delaysP[i], p_base$rho1, calcRe_exact(p_base)))
  }
}


out <- data.frame(out[-1,])
colnames(out) <- c("delayP", "rhoP", names(calcRe_exact(p_base)))

ggplot(out, aes(x=delayP, y=rhoP, fill=Re_I1)) +
  geom_tile() +
  geom_contour(aes(z=Re_I1, 
                   color = factor(..level.. == 1, levels = c(F, T), labels = c("", "Reff=1")))) + 
  scale_colour_manual(values = c("white", "red"), guide=F) +
  theme_bw()+
  scale_fill_viridis_c(name="R_eff") +
  xlab("Time from symptom onset to index isolation") +
  ylab("Proportion index cases detected")

```
