---
title: "What happened on the Diamond Princess?"
author: "Stephen A Lauer and Justin Lessler"
date: "2020-02-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r load}
library(tidyverse)
library(lubridate)
library(EpiEstim)
library(gridExtra)
library(doMC)
registerDoMC(detectCores()-2)
## color blind palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

## read in Diamond Princess data
dp_dat <- read_csv("data/diamond-princess.csv") %>% 
    mutate(dates=dmy(paste(Date, "2020", sep="-")),
           ## add the HK case to Tests and Positive
           Tests=ifelse(dates>=ymd("2020-02-01"), Tests+1, Tests),
           Positive=ifelse(dates>=ymd("2020-02-01"), Positive+1, Positive),
           Asymptomatic=ifelse(is.na(Asymptomatic),0,Asymptomatic),
           ## get daily numbers
           daily_test=diff(c(0,Tests)),
           daily_infected=diff(c(0,Positive)),
           daily_asym=diff(c(0,Asymptomatic)),
           daily_sym=daily_infected-daily_asym)
```

The Diamond Princess cruise was placed under quarantine on Feb 5 after a former passenger (who was on the boat from January 20-25) tested positive for COVID-19 in Hong Kong on Feb 1.
During the quarantine, which ended on Feb 19, a subset of passengers were tested for COVID-19 every day.
As infection numbers rose, many people questioned the efficacy of the quarantine.
On Feb 20, the Japanese National Institute of Infectious Diseases [reported](https://www.japantimes.co.jp/news/2020/02/20/national/coronavirus-diamond-princess-before-quarantine/) that the majority of transmission occurred prior to the quarantine.

Based on the publicly reported surveillance data, we estimated the daily reproduction number on the cruise, i.e. the number of people that each infected person was infecting each day.
A reproduction number above 1 indicates that the virus is spreading, while below 1 indicates that the virus is dying out.
We observe how estimates change when using the raw data and under a set of assumptions.

## Data overview

The Diamond Princess data consists of a dates and the cumulative tests, positive tests, and "asymptomatic" cases for those dates.
However, an "asymptomatic" case may merely be a case that has yet to develop symptoms, so that term should not be taken literally.
Asymptomatic cases are not formally counted until Feb 15, at which time it was noted that there were 35 previously discovered asymptomatic cases, which we assigned to Feb 13.
This results in more symptomatic cases at the beginning and more asymptomatic cases at the end.

Here is a summary of the testing protocol outlined in two field documents by NIID ([one](https://www.niid.go.jp/niid/en/2019-ncov-e/9407-covid-dp-fe-01.html) and [two](https://www.niid.go.jp/niid/en/2019-ncov-e/9417-covid-dp-fe-02.html)):

- Feb 3-4: All passengers observed, symptomatic individuals had swabs taken for testing.
- Feb 5: First test results come back, 1st positive COVID-19; quarantine starts
- Feb 7: Thermometers distributed, passengers told to self monitor
- Feb 11: Change in protocol, start testing by age group
- Feb 15: First day with “asymptomatic” positives recorded
- Feb 19: Start of disembarking for uninfected passengers

```{r data-plots}
data_plot <- dp_dat %>% 
    select(dates, daily_sym, daily_asym) %>% 
    pivot_longer(cols=c(daily_sym, daily_asym),
                 names_to="case_type",
                 values_to="number")

symptomatic_plot <- ggplot(data=data_plot) +
    geom_bar(aes(x=dates, y=number, fill=case_type),
             stat="identity", position="dodge") +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    geom_text(aes(x=ymd("2020-02-03"), y=70, label="Quarantine\nstarts"),
              color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Cases") +
    scale_fill_manual("Case type",
                      labels=c("Asymptomatic", "Symptomatic"),
                      values=cbbPalette[-1]) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.title.y=element_text(vjust=0),
          axis.text.x=element_blank(),
          legend.position=c(.15, .5),
          legend.background=element_blank())

test_plot <- ggplot(data=dp_dat) +
    geom_bar(aes(x=dates, y=daily_test),
             stat="identity", position="dodge") +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    geom_text(aes(x=ymd("2020-02-09"), y=500, label="Protocol\nchange"),
              color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Tests") +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.text.x=element_text(color="black", angle=45, hjust=1))

# grid.arrange(symptomatic_plot, test_plot, ncol=1)
```

## Using all cases

We use the `EpiEstim` package in R for our analysis, which requires dates, number of infections per date, and an estimate of the serial interval.
The serial interval is the time between the infections in a transmission chain, e.g. if I infect you with a disease, then it's the time between my infection and your infection.
This is frequently estimated by measuring the time between symptom onset times.
For this analysis, we use two serial interval estimates:

1. The estimate from [SARS](https://science.sciencemag.org/content/300/5627/1966) (mean: 8.4 days, sd: 3.8)
2. An early estimate for [COVID-19](https://www.medrxiv.org/content/10.1101/2020.02.03.20019497v2.full.pdf+html) (mean: 4.7, sd: 2.9)
<!-- 3. An eyeball estimate from our data (mean: 3.0, sd: 3.2) -->

We'll start out by looking at all of the data, including symptomatic and asymptomatic cases.

```{r Rt}
R_est_SARS <- estimate_R(dp_dat %>% select(dates, I=daily_infected),
                         method="parametric_si",
                         config=make_config(list(mean_si=8.4,
                                                 std_si=3.8,
                                                 # mean_si=4.41,
                                                 # std_si=3.17,
                                                 t_start=seq(2,nrow(dp_dat)),
                                                 t_end=seq(2,nrow(dp_dat)))))
# knitr::kable(R_est_SARS$R[,c(1,3,5,8,11)], digits=2)


R_est_ncov <- estimate_R(dp_dat %>% select(dates, I=daily_infected),
                         method="parametric_si",
                         config=make_config(list(#mean_si=8.4,
                             #std_si=3.8,
                             mean_si=4.4,
                             std_si=3.2,
                             t_start=seq(2,nrow(dp_dat)),
                             t_end=seq(2,nrow(dp_dat)))))

# knitr::kable(R_est_ncov$R[,c(1,3,5,8,11)], digits=2)

R_est_eyeball <- estimate_R(dp_dat %>% select(dates, I=daily_infected),
                            method="parametric_si",
                            config=make_config(list(#mean_si=8.4,
                                #std_si=3.8,
                                mean_si=3,
                                std_si=3.2,
                                t_start=seq(2,nrow(dp_dat)),
                                t_end=seq(2,nrow(dp_dat)))))

# knitr::kable(R_est_eyeball$R[,c(1,3,5,8,11)], digits=2)

plot_dat <- tibble(dates=c(R_est_SARS$dates[-1],
                           R_est_ncov$dates[-1],
                           R_est_eyeball$dates[-1]),
                   R_mean=c(R_est_SARS$R$`Mean(R)`,
                            R_est_ncov$R$`Mean(R)`,
                            R_est_eyeball$R$`Mean(R)`),
                   R_lb=c(R_est_SARS$R$`Quantile.0.025(R)`,
                          R_est_ncov$R$`Quantile.0.025(R)`,
                          R_est_eyeball$R$`Quantile.0.025(R)`),
                   R_ub=c(R_est_SARS$R$`Quantile.0.975(R)`,
                          R_est_ncov$R$`Quantile.0.975(R)`,
                          R_est_eyeball$R$`Quantile.0.975(R)`),
                   gen_time=c(rep("SARS", nrow(R_est_SARS$R)),
                              rep("COVID-19", nrow(R_est_ncov$R)),
                              rep("Eyeball", nrow(R_est_eyeball$R))),
                   analysis="overall") %>% 
    left_join(dp_dat %>% select(dates, daily_infected)) %>% 
    filter(daily_infected>0 | R_mean>0.2, gen_time !="Eyeball")

R0_plot <- ggplot(data=plot_dat, aes(x=dates, color=as.factor(gen_time))) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    geom_text(aes(x=ymd("2020-02-18"), y=32, label="US\nevacuation"),
              color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_line(aes(y=R_mean)) +
    # geom_line(aes(y=R_lb), linetype="dashed") +
    # geom_line(aes(y=R_ub), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7]) +
    theme_bw() +
    ggtitle("Estimates using all confirmed cases") +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position=c(0.15, 0.6),
          legend.background=element_blank())
# 
# ggplot(data=plot_dat, aes(x=dates, color=as.factor(gen_time))) +
#     geom_smooth(aes(y=R_mean), se=F) +
#     geom_smooth(aes(y=R_lb), linetype="dashed", se=F) +
#     geom_smooth(aes(y=R_ub), linetype="dashed", se=F) +
#     scale_x_date("", limits=c(ymd("2020-01-25"),ymd("2020-02-20"))) +
#     scale_y_log10("R") +
#     scale_color_discrete("Serial interval") +
#     theme_bw() +
#     theme(axis.text=element_text(color="black"),
#           legend.position=c(0.75, 0.75))


grid.arrange(R0_plot, symptomatic_plot, test_plot,
             ncol=1)
```

When using the SARS serial interval estimate, there is a large spike in the reproductive number right after the outbreak.
A reproduction number of 64 would certainly qualify as a super-spreading event, but there are reasons to be skeptical of using the SARS serial interval.
The serial interval is impacted by the contact rate; if a person infrequently interacts with others, the time between their symptom onset and their contacts will likely be longer.
On a cruise with swimming pools, buffets, and a [casino](https://www.princess.com/ships-and-experience/onboard-experience/activities/casino/), contact rates may be higher and thus the serial interval would be shorter.
We'll think of the SARS estimate as a worst-case scenario.

With the COVID-19 estimate, the reproductive number falls below 1 three days after the implementation of the quarantine (Feb 8), however this may be due to the fact that there was very little testing done on the 8th or 9th.
The reproductive number rebounded to 6 on Feb 10 before settling in between 2-3 until falling below 2 on Feb 18.

However, this analysis probably biases the reproduction value high.
The serial interval is often estimated as the time between the symptom onset of a primary case and the symptom onset of a secondary case.
Since these "asymptomatic" cases could be symptomatic cases that have not presented symptoms yet, then we are violating that definition.
By including more cases, we are inflating the amount of transmission and thus reproduction number would be biased high.

This also doesn't account for the delays between symptom onset and confirmation, which changed when the protocol changed.
All of the data appear to be confirmation dates, rather than the dates of symptom onset.
From what we can gather from the NIID reports and the surveillance data, there is a 1-3 day lag for data reported prior to Feb 11.
Then the protocol changes to testing by age group, starting with the eldest and febrile first.
Therefore, it is possible that a younger person could be symptomatic for a week prior to testing.

## Using only symptomatic cases, adjusted for symptom onset

In this analysis, we remove the asymptomatic cases and adjust the symptomatic cases.
We assume that the first case was reported at the symptom onset time.
Cases confirmed before Feb 11 are randomly assigned to have a symptom onset date 1-3 days prior to confirmation.
Cases confirmed after Feb 11 are randomly assigned to have a symptom onset date between Feb 10 and the day prior to confirmation.

```{r Rt-symp, cache=T}
# set.seed(1)
distribute_case_onsets <- function(dp_dat){
    date_vec <- ymd("2020-01-01")
    for(i in 1:nrow(dp_dat)){
        if(dp_dat$daily_sym[i]==0){
            next
        }
        tmp <- rep(dp_dat$dates[i], dp_dat$daily_sym[i])
        date_vec <- c(date_vec, tmp)
    }
    date_vec <- date_vec[-1]
    delays <- 0
    for(i in 2:length(date_vec)){
        if(date_vec[i] == ymd("2020-02-05"))
            delays[i] <- sample(2, 1)
        if(date_vec[i] == ymd("2020-02-06"))
            delays[i] <- sample(3, 1)
        if(date_vec[i] == ymd("2020-02-07")){
            if(i < 34){
                delays[i] <- sample(2:4, 1)
            } else
                delays[i] <- 1
        }
        if(date_vec[i] == ymd("2020-02-08"))
            delays[i] <- 2
        if(date_vec[i] == ymd("2020-02-09"))
            delays[i] <- 3
        if(date_vec[i] == ymd("2020-02-10"))
            delays[i] <- sample(3, 1)
        if(date_vec[i] > ymd("2020-02-10"))
            delays[i] <- sample(as.numeric(date_vec[i]-ymd("2020-02-10")),1)
    }
    new_dates <- date_vec-delays
    new_dat <- tibble(dates=seq(min(new_dates), max(new_dates), by="days")) %>% 
        left_join(tibble(dates=new_dates) %>% 
                      group_by(dates) %>% 
                      summarize(daily_shadow=n()), by="dates") %>% 
        full_join(dp_dat, by="dates") %>% 
        mutate(daily_shadow=ifelse(is.na(daily_shadow),
                                   0, daily_shadow)) %>% 
        arrange(dates)
    return(new_dat)
}

R_sims <- foreach(i=1:1000, .combine=rbind) %dopar%{
    # set.seed(i)
    tmp_dist <- distribute_case_onsets(dp_dat)
    tmp_R_SARS <- estimate_R(tmp_dist %>%
                            select(dates, I=daily_shadow),
                        method="parametric_si",
                        config=make_config(list(
                            mean_si=8.4,
                            std_si=3.8,
                            t_start=seq(2,nrow(tmp_dist)),
                            t_end=seq(2,nrow(tmp_dist)))))
    tmp_R_ncov <- estimate_R(tmp_dist %>%
                            select(dates, I=daily_shadow),
                        method="parametric_si",
                        config=make_config(list(
                            mean_si=4.4,
                            std_si=3.2,
                            t_start=seq(2,nrow(tmp_dist)),
                            t_end=seq(2,nrow(tmp_dist)))))
    return(tibble(sim=i,
                  dates=tmp_dist$dates[-1],
                  SARS_est=tmp_R_SARS$R$`Mean(R)`,
                  ncov_est=tmp_R_ncov$R$`Mean(R)`,
                  infections=tmp_dist$daily_shadow[-1]))
}
```

```{r Rt-symp-plots}
R0_shadow_dist_plot <- R_sims %>% 
    pivot_longer(cols=c(SARS_est, ncov_est),
                 names_to="type",
                 values_to="R_est") %>% 
    group_by(dates, type) %>% 
    summarise(R_mean=mean(R_est, na.rm=T),
              R_lb=quantile(R_est, probs=.025, na.rm=T),
              R_ub=quantile(R_est, probs=.975, na.rm=T)) %>% 
    ungroup() %>% 
    mutate(gray_out=if_else(type=="SARS_est",
                            dates>=ymd_hms("2020-02-05 00:00:00"),
                            dates>=ymd_hms("2020-02-08 00:00:00")),
           dates=as_datetime(dates)) %>% 
    bind_rows(R_sims %>% 
                  pivot_longer(cols=c(SARS_est, ncov_est),
                               names_to="type",
                               values_to="R_est") %>% 
                  group_by(dates, type) %>% 
                  summarise(R_mean=mean(R_est, na.rm=T),
                            R_lb=quantile(R_est, probs=.025, na.rm=T),
                            R_ub=quantile(R_est, probs=.975, na.rm=T)) %>% 
                  ungroup() %>% 
                  filter((type=="SARS_est" & dates==ymd("2020-02-05")) | 
                             (type=="ncov_est" & dates==ymd("2020-02-08"))) %>% 
                  mutate(dates=as_datetime(dates) - seconds(1),
                         gray_out=F)) %>% 
    ggplot(aes(x=dates, color=type, fill=type)) +
    geom_vline(aes(xintercept=ymd_hms("2020-02-05 00:00:00")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd_hms("2020-02-11 00:00:00")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd_hms("2020-02-16 00:00:00")), linetype="dashed") +
    geom_text(aes(x=ymd_hms("2020-02-13 00:00:00"), y=32,
                  label="Protocol\nchange"),
              color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_ribbon(aes(ymin=R_lb, ymax=R_ub, alpha=gray_out)) +
    geom_line(aes(y=R_mean, linetype=gray_out)) +
    scale_x_datetime("",
                 limits=c(ymd_hms("2020-01-25 00:00:00"),
                          ymd_hms("2020-02-20 23:59:00")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7],
                       labels=c("COVID-19", "SARS")) +
    scale_fill_manual("Serial interval",
                      values=cbbPalette[6:7],
                      labels=c("COVID-19", "SARS")) +
    scale_alpha_manual(values=c(0.5,0.1),
                       guide=F) +
    scale_linetype_discrete(guide=F) +
    theme_bw() +
    ggtitle("Estimates with allocated symptom onsets, asymptomatics removed") +
    coord_cartesian(ylim=c(0.25, 64)) +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position=c(0.1, 0.6),
          legend.background=element_blank())

dist_inf_plot <- R_sims %>% 
    group_by(dates) %>% 
    summarise(inf_mean=mean(infections, na.rm=T),
              inf_lb=quantile(infections, probs=.025, na.rm=T),
              inf_ub=quantile(infections, probs=.975, na.rm=T)) %>% 
    mutate(gray_out=dates >= ymd("2020-02-13")) %>% 
    ggplot(aes(x=dates, y=inf_mean, alpha=as.factor(gray_out))) +
    geom_bar(stat="identity", position="dodge") +
    geom_errorbar(aes(ymin=inf_lb, ymax=inf_ub)) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    geom_text(aes(x=ymd("2020-02-03"), y=40, label="Quarantine\nstarts"),
              color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    geom_text(aes(x=ymd("2020-02-18"), y=40, label="US\nevacuation"),
              color="black") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Average cases, reallocated") +
    scale_alpha_manual(values=c(1,0.5),
                       guide=F) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.title.y=element_text(vjust=0),
          axis.text.x=element_text(color="black", angle=45, hjust=1),
          legend.position=c(.1, .5),
          legend.background=element_blank())

grid.arrange(R0_shadow_dist_plot, dist_inf_plot, ncol=1)
```

```{r old, eval=F}
R_SARS_sym <- estimate_R(dp_dat %>% select(dates, I=daily_sym),
                         method="parametric_si",
                         config=make_config(list(mean_si=8.4,
                                                 std_si=3.8,
                                                 t_start=seq(2,nrow(dp_dat)),
                                                 t_end=seq(2,nrow(dp_dat)))))


R_ncov_sym <- estimate_R(dp_dat %>% select(dates, I=daily_sym),
                         method="parametric_si",
                         config=make_config(list(#mean_si=8.4,
                             #std_si=3.8,
                             mean_si=4.4,
                             std_si=3.2,
                             t_start=seq(2,nrow(dp_dat)),
                             t_end=seq(2,nrow(dp_dat)))))

# knitr::kable(R_est_ncov$R[,c(1,3,5,8,11)], digits=2)

R_eyeball_sym <- estimate_R(dp_dat %>% select(dates, I=daily_sym),
                            method="parametric_si",
                            config=make_config(list(#mean_si=8.4,
                                #std_si=3.8,
                                mean_si=3,
                                std_si=3.2,
                                t_start=seq(2,nrow(dp_dat)),
                                t_end=seq(2,nrow(dp_dat)))))

# knitr::kable(R_est_eyeball$R[,c(1,3,5,8,11)], digits=2)

sym_plot_dat <- tibble(dates=c(R_SARS_sym$dates[-1],
                           R_ncov_sym$dates[-1],
                           R_eyeball_sym$dates[-1]),
                   R_mean=c(R_SARS_sym$R$`Mean(R)`,
                            R_ncov_sym$R$`Mean(R)`,
                            R_eyeball_sym$R$`Mean(R)`),
                   R_lb=c(R_SARS_sym$R$`Quantile.0.025(R)`,
                          R_ncov_sym$R$`Quantile.0.025(R)`,
                          R_eyeball_sym$R$`Quantile.0.025(R)`),
                   R_ub=c(R_SARS_sym$R$`Quantile.0.975(R)`,
                          R_ncov_sym$R$`Quantile.0.975(R)`,
                          R_eyeball_sym$R$`Quantile.0.975(R)`),
                   gen_time=c(rep("SARS", nrow(R_SARS_sym$R)),
                              rep("COVID-19", nrow(R_ncov_sym$R)),
                              rep("Eyeball", nrow(R_eyeball_sym$R))),
                   analysis="symptomatic") %>% 
    left_join(dp_dat %>% select(dates, daily_infected)) %>% 
    filter(daily_infected>0 | R_mean>0.2, gen_time !="Eyeball")

R0_sym_plot <- ggplot(data=sym_plot_dat, aes(x=dates, color=as.factor(gen_time))) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=100, label="US\nevacuation"),
    #           color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_line(aes(y=R_mean)) +
    # geom_line(aes(y=R_lb), linetype="dashed") +
    # geom_line(aes(y=R_ub), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7]) +
    theme_bw() +
    ggtitle("Estimates using all symptomatic cases") +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position="top")

symptomatic_only_plot <- data_plot %>% filter(case_type=="daily_sym") %>% 
    ggplot() +
    geom_bar(aes(x=dates, y=number),
             stat="identity", position="dodge") +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=70, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=70, label="US\nevacuation"),
    #           color="black") +
    scale_x_date("",
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Symptomatic cases") +
    scale_fill_manual("Case type",
                      labels=c("Asymptomatic", "Symptomatic"),
                      values=cbbPalette[-1]) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.title.y=element_text(vjust=0),
          axis.text.x=element_text(color="black", angle=45, hjust=1),
          legend.position=c(.15, .5),
          legend.background=element_blank())


R0_compare_plot <- bind_rows(plot_dat, sym_plot_dat) %>%
    filter(daily_infected>0 | R_mean>0.2) %>% 
    ggplot(aes(x=dates, color=gen_time, linetype=analysis)) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=100, label="US\nevacuation"),
    #           color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_line(aes(y=R_mean)) +
    # geom_line(aes(y=R_lb), linetype="dashed") +
    # geom_line(aes(y=R_ub), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7]) +
    scale_linetype_discrete("Analysis") +
    theme_bw() +
    ggtitle("Comparison of estimates with and without asymptomatic cases") +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position="top")

grid.arrange(R0_compare_plot, symptomatic_only_plot, ncol=1, layout_matrix=rbind(1, 1, 1, 2, 2))
```

<!-- Since the asymptomatic cases are concentrated at the end, that's where we see the difference in the estimates. -->
<!-- While we still observe the bounce on Feb 10, the estimates call below 1 again on Feb 13 and the average stays below 2 for the remainder of the crisis. -->

<!-- Of course, when we leave out asymptomatic cases and truncate the data on Feb 20 then we are ignoring a large portion of the transmission. -->
<!-- Let's see if we can account for these cases in a more intelligent manner. -->

<!-- ## Putting symptomatic cases on asymptomatic timeline -->

<!-- Instead of treating the asymptomatic cases as symptomatic or not, we could use adjust symptomatic cases earlier so that they are comparable to asymptomatic cases. -->
<!-- Here we reallocate the symptomatic cases based on their incubation period, which is the time between exposure to infection and symptom onset. -->
<!-- In an [earlier paper](https://www.medrxiv.org/content/10.1101/2020.02.02.20020016v1), we estimated that the median incubation time for COVID-19 is 5.2 days. -->
<!-- Thus, we first try moving symptomatic cases 5 days earlier: -->

```{r Rt-shadow, eval=F}
shadow_dat <- dp_dat %>% 
    ## prep scenario where asymptomatics pushed back 5 days
    mutate(daily_shadow=lead(daily_sym, 5, default=0),
           daily_infect_shadow=daily_asym+daily_shadow)

R_SARS_shadow <- estimate_R(shadow_dat %>% select(dates, I=daily_infect_shadow),
                            method="parametric_si",
                            config=make_config(list(mean_si=8.4,
                                                    std_si=3.8,
                                                    # mean_si=4.41,
                                                    # std_si=3.17,
                                                    t_start=seq(2,nrow(shadow_dat)),
                                                    t_end=seq(2,nrow(shadow_dat)))))
# knitr::kable(R_SARS_shadow$R[,c(1,3,5,8,11)], digits=2)


R_ncov_shadow <- estimate_R(shadow_dat %>% select(dates, I=daily_infect_shadow),
                            method="parametric_si",
                            config=make_config(list(
                                mean_si=4.4,
                                std_si=3.2,
                                t_start=seq(2,nrow(shadow_dat)),
                                t_end=seq(2,nrow(shadow_dat)))))

# knitr::kable(R_ncov_shadow$R[,c(1,3,5,8,11)], digits=2)

R_eyeball_shadow <- estimate_R(shadow_dat %>% select(dates, I=daily_infect_shadow),
                               method="parametric_si",
                               config=make_config(list(
                                   mean_si=3,
                                   std_si=3.2,
                                   t_start=seq(2,nrow(shadow_dat)),
                                   t_end=seq(2,nrow(shadow_dat)))))

# knitr::kable(R_eyeball_shadow$R[,c(1,3,5,8,11)], digits=2)

shadow_plot_dat <- tibble(dates=c(R_SARS_shadow$dates[-1],
                                  R_ncov_shadow$dates[-1],
                                  R_eyeball_shadow$dates[-1]),
                          R_mean=c(R_SARS_shadow$R$`Mean(R)`,
                                   R_ncov_shadow$R$`Mean(R)`,
                                   R_eyeball_shadow$R$`Mean(R)`),
                          R_lb=c(R_SARS_shadow$R$`Quantile.0.025(R)`,
                                 R_ncov_shadow$R$`Quantile.0.025(R)`,
                                 R_eyeball_shadow$R$`Quantile.0.025(R)`),
                          R_ub=c(R_SARS_shadow$R$`Quantile.0.975(R)`,
                                 R_ncov_shadow$R$`Quantile.0.975(R)`,
                                 R_eyeball_shadow$R$`Quantile.0.975(R)`),
                          gen_time=c(rep("SARS", nrow(R_SARS_shadow$R)),
                                     rep("COVID-19", nrow(R_ncov_shadow$R)),
                                     rep("Eyeball", nrow(R_eyeball_shadow$R))),
                          analysis="early") %>% 
    left_join(shadow_dat %>% select(dates, daily_infect_shadow)) %>% 
    filter(daily_infect_shadow>0 | R_mean>0.2, gen_time !="Eyeball")

R0_shadow_plot <- ggplot(data=shadow_plot_dat,
                         aes(x=dates, color=as.factor(gen_time))) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=100, label="US\nevacuation"),
    #           color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_line(aes(y=R_mean)) +
    # geom_line(aes(y=R_lb), linetype="dashed") +
    # geom_line(aes(y=R_ub), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7]) +
    theme_bw() +
    ggtitle("Estimates on asymptomatic timeline") +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position=c(0.15, 0.6),
          legend.background=element_blank())

asymptomatic_tl_plot <- ggplot(data=shadow_dat) +
    geom_bar(aes(x=dates, y=daily_infect_shadow),
             stat="identity", position="dodge") +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=70, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=70, label="US\nevacuation"),
    #           color="black") +
    scale_x_date("",
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Reallocated cases") +
    scale_fill_manual("Case type",
                      labels=c("Asymptomatic", "Symptomatic"),
                      values=cbbPalette[-1]) +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.title.y=element_text(vjust=0),
          axis.text.x=element_text(color="black", angle=45, hjust=1),
          legend.position=c(.15, .5),
          legend.background=element_blank())



# grid.arrange(R0_plot, symptomatic_plot, test_plot, ncol=1)

R0_compare_shadow_plot <- bind_rows(plot_dat,
          # sym_plot_dat,
          shadow_plot_dat) %>%
    filter(daily_infected>0 | R_mean>0.2, gen_time=="COVID-19") %>% 
    ggplot(aes(x=dates, color=analysis)) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
    #           color="black") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    # geom_text(aes(x=ymd("2020-02-14"), y=100, label="US\nevacuation"),
    #           color="black") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_line(aes(y=R_mean)) +
    # geom_line(aes(y=R_lb), linetype="dashed") +
    # geom_line(aes(y=R_ub), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Analysis",
                       values=cbbPalette[c(7:6)]) +
    theme_bw() +
    ggtitle("Comparison of reallocated estimates using COVID-19 serial intervals") +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position=c(0.1, 0.6),
          legend.background=element_blank())

grid.arrange(R0_compare_shadow_plot, asymptomatic_tl_plot, ncol=1)
```

<!-- Moving the symptomatic cases 5 days earlier, results in an earlier peak in R, as we would expect. -->
<!-- The values for each day up to Feb 12 in the original analysis are shifted 5 days early. -->
<!-- Out of luck or circumstance, there again were very few confirmed cases on Feb 8 and 9 and the estimate falls below 1. -->
<!-- However, instead of rebounding to 6, the estimates stay below 3 for the rest of the way. -->

<!-- Instead of moving every symptomatic case early by a constant amount, we can reallocate the incubation period according to a distribution. -->
<!-- This makes sense because incubation period vary in length from person to person, with some people having incubations as short as a day and others having incubation periods as long as 2-3 weeks. -->
<!-- We apply this to all infections, symptomatic and asymptomatic; otherwise we'd be assuming that all of the asymptomatic infections had just begun, when they may have been going on for a while. -->

```{r Rt-shadow-dist, eval=F}
# set.seed(1)
distribute_case_delays <- function(dp_dat){
    date_vec <- ymd("2020-01-01")
    for(i in 1:nrow(dp_dat)){
        if(dp_dat$daily_infected[i]==0){
            next
        }
        tmp <- rep(dp_dat$dates[i], dp_dat$daily_infected[i])
        date_vec <- c(date_vec, tmp)
    }
    date_vec <- date_vec[-1]
    delays <- rlnorm(length(date_vec), 1.63, 0.36) %>% round()
    new_dates <- pmax(date_vec-delays, ymd("2020-01-20"))
    new_dat <- tibble(dates=seq(min(new_dates), max(new_dates), by="days")) %>% 
        left_join(tibble(dates=new_dates) %>% 
                      group_by(dates) %>% 
                      summarize(daily_shadow=n()), by="dates") %>% 
        full_join(dp_dat, by="dates") %>% 
        mutate(daily_shadow=ifelse(is.na(daily_shadow),
                                   0, daily_shadow),
               daily_asym=ifelse(is.na(daily_asym),
                                   0, daily_asym),
               daily_infect_shadow=daily_asym+daily_shadow)
    return(new_dat)
}
# shadow_dat_dist <- distribute_case_delays(dp_dat)
# 
# shadow_dat_dist %>% 
#     select(dates, original=daily_infected, simulated=daily_infect_shadow) %>% 
#     pivot_longer(cols=c(original, simulated),
#                  names_to="case_type",
#                  values_to="number") %>% 
#     ggplot() +
#     geom_bar(aes(x=dates, y=number, fill=case_type),
#              stat="identity", position="dodge") +
#     geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
#     # geom_text(aes(x=ymd("2020-02-03"), y=70, label="Quarantine\nstarts"),
#     #           color="black") +
#     geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
#     # geom_text(aes(x=ymd("2020-02-14"), y=70, label="US\nevacuation"),
#     #           color="black") +
#     scale_x_date("",
#                  date_breaks="2 days",
#                  date_labels="%b %d") +
#     scale_y_continuous("Number of cases") +
#     scale_fill_manual("",
#                       labels=c("Original", "Simulated"),
#                       values=cbbPalette[-1]) +
#     theme_bw() +
#     theme(axis.text.y=element_text(color="black"),
#           axis.title.x=element_blank(),
#           axis.title.y=element_text(vjust=0),
#           axis.text.x=element_blank(),
#           legend.position=c(.15, .5))
# R_SARS_shadow_dist <- estimate_R(shadow_dat_dist %>%
#                                      select(dates, I=daily_infect_shadow),
#                             method="parametric_si",
#                             config=make_config(
#                                 list(mean_si=8.4,
#                                      std_si=3.8,
#                                      t_start=seq(2,nrow(shadow_dat_dist)),
#                                      t_end=seq(2,nrow(shadow_dat_dist)))))
# # knitr::kable(R_SARS_shadow_dist$R[,c(1,3,5,8,11)], digits=2)
# 
# 
# R_ncov_shadow_dist <- estimate_R(shadow_dat_dist %>%
#                                      select(dates, I=daily_infect_shadow),
#                             method="parametric_si",
#                             config=make_config(list(
#                                 mean_si=4.4,
#                                 std_si=3.2,
#                                 t_start=seq(2,nrow(shadow_dat_dist)),
#                                 t_end=seq(2,nrow(shadow_dat_dist)))))
# 
# # knitr::kable(R_ncov_shadow_dist$R[,c(1,3,5,8,11)], digits=2)
# 
# R_eyeball_shadow_dist <- estimate_R(shadow_dat_dist %>%
#                                         select(dates, I=daily_infect_shadow),
#                                method="parametric_si",
#                                config=make_config(list(
#                                    mean_si=3,
#                                    std_si=3.2,
#                                    t_start=seq(2,nrow(shadow_dat_dist)),
#                                    t_end=seq(2,nrow(shadow_dat_dist)))))
# 
# # knitr::kable(R_eyeball_shadow_dist$R[,c(1,3,5,8,11)], digits=2)
# 
# shadow_plot_dat_dist <- tibble(dates=c(R_SARS_shadow_dist$dates[-1],
#                                   R_ncov_shadow_dist$dates[-1],
#                                   R_eyeball_shadow_dist$dates[-1]),
#                           R_mean=c(R_SARS_shadow_dist$R$`Mean(R)`,
#                                    R_ncov_shadow_dist$R$`Mean(R)`,
#                                    R_eyeball_shadow_dist$R$`Mean(R)`),
#                           R_lb=c(R_SARS_shadow_dist$R$`Quantile.0.025(R)`,
#                                  R_ncov_shadow_dist$R$`Quantile.0.025(R)`,
#                                  R_eyeball_shadow_dist$R$`Quantile.0.025(R)`),
#                           R_ub=c(R_SARS_shadow_dist$R$`Quantile.0.975(R)`,
#                                  R_ncov_shadow_dist$R$`Quantile.0.975(R)`,
#                                  R_eyeball_shadow_dist$R$`Quantile.0.975(R)`),
#                           gen_time=c(rep("SARS", nrow(R_SARS_shadow_dist$R)),
#                                      rep("COVID-19", nrow(R_ncov_shadow_dist$R)),
#                                      rep("Eyeball", nrow(R_eyeball_shadow_dist$R))),
#                           analysis="early") %>% 
#     left_join(shadow_dat_dist %>% select(dates, daily_infect_shadow)) %>% 
#     filter(daily_infect_shadow>0 | R_mean>0.2, gen_time !="Eyeball")
# 
# ggplot(data=shadow_plot_dat_dist, aes(x=dates, color=as.factor(gen_time))) +
#     geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
#     # geom_text(aes(x=ymd("2020-02-03"), y=100, label="Quarantine\nstarts"),
#     #           color="black") +
#     geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
#     # geom_text(aes(x=ymd("2020-02-14"), y=100, label="US\nevacuation"),
#     #           color="black") +
#     geom_hline(aes(yintercept=1), linetype="dashed") +
#     geom_line(aes(y=R_mean)) +
#     # geom_line(aes(y=R_lb), linetype="dashed") +
#     # geom_line(aes(y=R_ub), linetype="dashed") +
#     scale_x_date("",
#                  # limits=c(ymd("2020-01-25"),ymd("2020-02-20")),
#                  date_breaks="2 days",
#                  date_labels="%b %d") +
#     scale_y_log10("Time-varying reproduction number, R",
#                   breaks=2^seq(-10,10,2)) +
#     scale_color_manual("Serial interval",
#                        values=cbbPalette[-1]) +
#     theme_bw() +
#     theme(axis.text.y=element_text(color="black"),
#           axis.text.x=element_text(color="black", angle=45, hjust=1),
#           legend.position="bottom")
# # 
# # ggplot(data=plot_dat, aes(x=dates, color=as.factor(gen_time))) +
# #     geom_smooth(aes(y=R_mean), se=F) +
# #     geom_smooth(aes(y=R_lb), linetype="dashed", se=F) +
# #     geom_smooth(aes(y=R_ub), linetype="dashed", se=F) +
# #     scale_x_date("", limits=c(ymd("2020-01-25"),ymd("2020-02-20"))) +
# #     scale_y_log10("R") +
# #     scale_color_discrete("Serial interval") +
# #     theme_bw() +
# #     theme(axis.text=element_text(color="black"),
# #           legend.position=c(0.75, 0.75))
# 
# 
# # grid.arrange(R0_plot, symptomatic_plot, test_plot, ncol=1)
# 
# bind_rows(plot_dat,
#           # sym_plot_dat,
#           shadow_plot_dat_dist) %>%
#     filter(daily_infected>0 | R_mean>0.2, gen_time=="COVID-19") %>% 
#     ggplot(aes(x=dates, color=analysis)) +
#     geom_line(aes(y=R_mean)) +
#     # geom_line(aes(y=R_lb), linetype="dashed") +
#     # geom_line(aes(y=R_ub), linetype="dashed") +
#     geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
#     geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
#     geom_hline(aes(yintercept=1), linetype="dashed") +
#     scale_x_date("",
#                  date_breaks="2 days",
#                  date_labels="%b %d") +
#     scale_y_log10("Time-varying reproduction number, R",
#                   breaks=2^seq(-10,10,2)) +
#     scale_color_manual("Analysis",
#                        breaks=c("overall", "symptomatic", "early"),
#                        values=cbbPalette[c(8, 6, 7)]) +
#     ggtitle("Comparison of analyses using COVID-19 estimates") +
#     theme_bw() +
#     theme(axis.text.y=element_text(color="black"),
#           axis.text.x=element_text(color="black", angle=45, hjust=1),
#           legend.position="bottom")
```

<!-- We simulate this process 1,000 times and take the average and 95% confidence interval across all simulations: -->

```{r Rt-shadow-dist-1000, eval=F}
R_sims <- foreach(i=1:1000, .combine=rbind) %dopar%{
    # set.seed(i)
    tmp_dist <- distribute_case_delays(dp_dat)
    tmp_R <- estimate_R(tmp_dist %>%
                            select(dates, I=daily_infect_shadow),
                        method="parametric_si",
                        config=make_config(list(
                            mean_si=4.4,
                            std_si=3.2,
                            t_start=seq(2,nrow(tmp_dist)),
                            t_end=seq(2,nrow(tmp_dist)))))
    return(tibble(sim=i,
                  dates=tmp_R$dates[-1],
                  R_est=tmp_R$R$`Mean(R)`,
                  infections=tmp_dist$daily_infect_shadow[-1]))
}

R0_shadow_dist_plot <- R_sims %>% 
    group_by(dates) %>% 
    summarise(obs=sum(!is.na(R_est)),
              R_mean=mean(R_est, na.rm=T),
              R_lb=quantile(R_est, probs=.025, na.rm=T),
              R_ub=quantile(R_est, probs=.975, na.rm=T)) %>% 
    ggplot(aes(x=dates, y=R_mean)) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    geom_hline(aes(yintercept=1), linetype="dashed") +
    geom_ribbon(aes(ymin=R_lb, ymax=R_ub), alpha=0.4) +
    geom_line(aes(y=R_mean)) +
    scale_x_date("",
                 limits=c(ymd("2020-01-21"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_log10("R, log scale",
                  breaks=2^seq(-2,6,2),
                  labels=c("1/4", "1", "4", "16", "64")) +
    scale_color_manual("Serial interval",
                       values=cbbPalette[6:7]) +
    theme_bw() +
    ggtitle("Estimates with distributed incubation period") +
    coord_cartesian(ylim=c(0.25, 16)) +
    theme(axis.text.y=element_text(color="black"),
          # axis.text.x=element_text(color="black", angle=45, hjust=1),
          axis.text.x=element_blank(),
          axis.title.x=element_blank(),
          legend.position=c(0.15, 0.6),
          legend.background=element_blank())

dist_inf_plot <- R_sims %>% 
    group_by(dates) %>% 
    summarise(obs=sum(!is.na(R_est)),
              inf_mean=mean(infections, na.rm=T),
              inf_lb=quantile(infections, probs=.025, na.rm=T),
              inf_ub=quantile(infections, probs=.975, na.rm=T)) %>% 
    ggplot(aes(x=dates, y=inf_mean)) +
    geom_bar(stat="identity", position="dodge") +
    geom_errorbar(aes(ymin=inf_lb, ymax=inf_ub)) +
    geom_vline(aes(xintercept=ymd("2020-02-05")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-11")), linetype="dashed") +
    geom_vline(aes(xintercept=ymd("2020-02-16")), linetype="dashed") +
    scale_x_date("",
                 limits=c(ymd("2020-01-21"),ymd("2020-02-20")),
                 date_breaks="2 days",
                 date_labels="%b %d") +
    scale_y_continuous("Average reallocated cases") +
    theme_bw() +
    theme(axis.text.y=element_text(color="black"),
          axis.title.x=element_blank(),
          axis.title.y=element_text(vjust=0),
          axis.text.x=element_text(color="black", angle=45, hjust=1),
          legend.position=c(.15, .5),
          legend.background=element_blank())

grid.arrange(R0_shadow_dist_plot, dist_inf_plot, ncol=1)
```

The faded lines and bars show where there is missing asymptomatic data, which would affect those estimates.

Under these assumptions, we see that the peak reproduction number (and thus transmission) may have occurred prior to quarantine and that the quarantine may have helped in reducing the spread of COVID-19.
However, even without the asymptomatic cases, the reproduction number doesn't fall below 1 until Feb 12 at the earliest.

As more information becomes available we will update this analysis.
