---
title: "Simple CFR approach"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

 
require(knitr)
require(tidyverse)
require(gridExtra)
require(rstan)
require(splines2)
require(readr)


source("../R/DataLoadUtils.r")
source("../R/BasicEpiAnalyses.r")
source("../R/CFRutils.r")

```

## Goal
Here is the basic model. We are interested in estimating the CFR among confirmed 
cases, $\rho$. 

Presume that the time from confirmation to report of death (censoring at recovery) follows an arbitrary parametric distribution:
$$\Pr(Y_i<k) = F(k; \theta)$$

Where $Y_i$ is the time from confirmation to  death for person $i$ and $\theta$ is a vector of parameters required to define the parametric distribution..

Moreover, the time from confirmation to death or recovery (composite outcome) follows an exponential distribution:
$$\Pr(T_i<k) = G(k; \eta)$$
Where $T_i$ is the time from confirmation to  death or recovery for person $i$.



Let $d_{jt}$ be the number of new reported deaths in location $j$ on calendar date $t$. $d_{jt}$ depends on the hazard function for death over time since becoming a confirmed case and the distribution of infection duration on calendar date $t$.
$$E(d_{jt}) = \sum_{k=1}^t n_{jtk}h_{1k}$$

where $h_{1k}$ is the cause-specific hazard function for death over time since confirmation, and whose shape is determined by parameters $\theta$.


Let $r_{jt}$ be the number of new reported recoveries in location $j$ on calendar date $t$. $r_{jt}$ depends on the hazard function for recovery over time since becoming a confirmed case and the distribution of infection duration on calendar date $t$.
$$E(r_{jt}) = \sum_{k=1}^t n_{jtk}h_{2k}$$

where $h_{2k}$ is the cause-specific hazard function for recovery over time since confirmation, and whose shape is determined by parameters $\eta$.

For example, if we assume that $F{k}$ follows a Weibull distribution, $\theta = \{\lambda_1, \alpha_1\}$ and $h_{1k} = \alpha_1\lambda_1^{\alpha_1}k^{(\alpha_1-1)}$

To simplify estimation of $\rho$ in our dataset, which is organized by calendar date, we will first assume both time to death and time to the composite outcome follow an exponential distirbution with a constant hazard functions, $\lambda_1$ and $\lambda$, respectively, for now. 

Based on the above:
$$E(d_{jt}) =  n_{jt} \lambda_1$$
where $n_{jt}$ is the number of currently infected people in the risk set on that day.
Similarly, let $r_{jt}$ be the number of new recoveries s in location $j$ on calendar date $t$. Based on the 
above:
$$E(r_{jt}) =  n_{jt} (\lambda - \lambda_1) $$

## Stan models
First, define the stan model.

```{stan, eval = FALSE, output.var="cfrmdl2"}

data {
  int <lower=0> T; //the number of time steps included
  int <lower=0> L; //the number of locations we have data from
  int <lower=0> r[T,L]; //  number of recovered cases reported on each day.
  int <lower=0> d[T,L]; //  number of deaths reported on each day. 
  real <lower=0> c[T,L]; //number of new cases  on each day. 
  int w[L]; //is this wuhan
}

parameters {
  real <lower=0> lambda1; //parameter for time to death distribution
  real <lower=0> lambda; // parameter for time to death or recovery distribution
  real <lower= 0, upper=1> dr; // wuhan detect rate
}

transformed parameters {
  real <lower=0> expected_deaths[T,L]; //  # deaths at time T
  real <lower=0> expected_recovereds[T,L]; //  # those who have exited due to recovery at T
  real <lower=0> expected_riskset[T,L]; // expected people in risk set

for (j in 1:L) {
    expected_riskset[1,j] = c[1,j];
    expected_deaths[1,j] = expected_riskset[1,j] * lambda1 + 0.0001 ;
    expected_recovereds[1,j] = expected_riskset[1,j] * (lambda - lambda1) * (dr)^w[j] + 0.0001;
 
    for (t in 2:T) {
      expected_riskset[t,j] =  expected_riskset[t-1,j] + c[t,j] - 
                                expected_deaths[t-1,j] - expected_riskset[t-1,j] * (lambda - lambda1);
      expected_deaths[t,j] = expected_riskset[t,j] * lambda1 + 0.0001 ;
      expected_recovereds[t,j] = expected_riskset[t,j] * (lambda - lambda1) * (dr)^w[j] + 0.0001;
      
    }
  }
}


model {
  //definitely can be made more effcient.
  for (j in 1:L) {
    for (t in 1:T) {
      target+=poisson_lpmf(d[t,j]|expected_deaths[t,j]) + poisson_lpmf(r[t,j]|expected_recovereds[t,j]);
    }
  }
 
}


```

Now, attempt to allow hazards to vary over time (this model is under construction).
```{stan, eval = FALSE, output.var="cfrmdl3b"}

data {
  int <lower=0> T; //the number of time steps included
  int <lower=0> L; //the number of locations we have data from
  int <lower = 0> V; // max infection duration (may not be needed)
  int <lower=0> r[T,L]; //  number of recovered cases reported on each day.
  int <lower=0> d[T,L]; //  number of deaths reported on each day. 
  real <lower=0> c[T,L]; //number of new cases  on each day. 
  int w[L]; //is this wuhan
}

parameters {
  real loglambda1; //parameter for time to death distribution
  real loglambda2; // parameter for time to death or recovery distribution
  real  logalpha1; //parameter for time to death distribution
  real  logalpha2; // parameter for time to death or recovery distribution
  real  logitdr; // wuhan detect rate
}
transformed parameters {
  real <lower=0> expected_deaths[T,L]; //  # deaths at time T
  real <lower=0> expected_recovereds[T,L]; //  # those who have exited due to recovery at T
  //matrix[V,T] expected_riskset[L]; // expected people in risk set in inf dur timescale
  real <lower=0> expected_riskset[V,T]; //expected risk set on each day of inf dur
  real <lower=0> h1[V];
  real <lower=0> h2[V];
  real <lower=0> lambda1; //parameter for time to death distribution
  real <lower=0> lambda2; // parameter for time to death or recovery distribution
  real <lower=0> alpha1; //parameter for time to death distribution
  real <lower=0> alpha2; // parameter for time to death or recovery distribution
  real <lower=0, upper=1> dr; //detectino rate

  lambda1 = exp(loglambda1);
  lambda2 = exp(loglambda2);
  alpha1 = exp(logalpha1);
  alpha2 = exp(logalpha2);
  dr = 1/(1+exp(-logitdr));

for (j in 1:L) {
   // expected_riskset[j][1,1] = c[1,j]+0.0001; //risk set on day 1 at inf duration 1 is just c1
   expected_riskset[1,1] = c[1,j]+0.0001; //risk set on day 1 at inf duration 1 is just c1
    expected_deaths[1,j] = expected_riskset[1,1] * alpha1*lambda1^alpha1 + 0.0001 ;
    expected_recovereds[1,j] = expected_riskset[1,1] * alpha2*lambda2^alpha2 * (dr)^w[j] + 0.0001;
 

    for (t in 2:T) {
      expected_riskset[1,t] = c[t,j]+0.0001;
      expected_deaths[t,j] = expected_riskset[1,t] * alpha1*lambda1^alpha1 + 0.0001 ;
      expected_recovereds[t,j] = expected_riskset[1,t] * alpha2*lambda2^alpha2 * (dr)^w[j] + 0.0001;
      h1[1] = alpha1*lambda1^alpha1;
      h2[1] = alpha2*lambda2^alpha2;
      
      for(k in 2:t){
       h1[k] = alpha1*lambda1^alpha1*k^(alpha1-1);
       h2[k] = alpha2*lambda2^alpha2*k^(alpha2-1);
        expected_riskset[k,t] = (expected_riskset[k-1,t-1] - (expected_riskset[k-1,t-1]*h1[k-1] +  (expected_riskset[k-1,t-1]* h2[k-1]))+0.0001);
        expected_deaths[t,j] +=  ((expected_riskset[k,t] * h1[k]) +0.0001;
        expected_recovereds[t,j] += ((expected_riskset[k,t] * h2[k] * (dr)^w[j]) +0.0001;

      }
    }
  }
}


model {
  //definitely can be made more effcient.
  for (j in 1:L) {
    for (t in 1:T) {
      target+=poisson_lpmf(d[t,j]|expected_deaths[t,j]) + poisson_lpmf(r[t,j]|expected_recovereds[t,j]);
    }
  }
 
}

```

## Data management
Prep data 

```{r message=FALSE, warning=FALSE, include=FALSE}
library(readr)
library(tidyverse)
   #Load in the JHU CSSE Data
todaydate <- ISOdate(2020,2,13, hour = 00, min = 54, tz = "UTC")
#note, weird stuff going on with the dates. so i took the max time in the newest file and inserted here, which fixed.
date <- todaydate

#mod2_nohubei <- moddata(todaydate, outsideChina = FALSE, ChinaOnly = TRUE, withHubei = FALSE, HubeiOnly = FALSE)
mod2_all <- moddata(todaydate, withHubei = TRUE)
#mod2_hubei <- moddata(todaydate, outsideChina = FALSE, ChinaOnly = TRUE, withHubei = TRUE, HubeiOnly = TRUE)
#mod2_nochina <- moddata(todaydate, outsideChina = TRUE)


```

## Run models

```{r, eval = FALSE}
  cfrmdl_je_res <- sampling(cfrmdl2, data=mod2_all,
                         iter=3000)
  chains <- extract(cfrmdl_je_res)
  lambda1 <- chains$lambda1
  #lambda2 <- chains$lambda2
  lambda <- chains$lambda
  # alpha1 <- chains$alpha1
  #alpha2 <- chains$alpha2
  dr <- chains$dr
  mean(dr)
  #outdat <- data.frame(lambda1 = lambda1, lambda2 = lambda, alpha1 = alpha1, alpha2 = alpha2 , dr = dr)
  outdat <- data.frame(lambda1 = lambda1, lambda = lambda, dr = dr)
  save(outdat, file = "../data/outdat.RData")

```

Summarize


```{r, echo = FALSE }
load("../data/outdat.RData")

lambda1 <- outdat[,1] #chains$lambda1
lambda <- outdat[,2] #chains$lambda
#alpha1 <- outdat[,3]
#alpha2 <- outdat[,4]
dr <- outdat[,3]
  
plot(density(lambda1), xlim=c(0,.1), type = "l")
 lines(density(lambda))
 plot(lambda, type = "l")
 plot(lambda1, type = "l")



```

Results
```{r}

#move this into stan model later
rho <- matrix(nrow = length(lambda1))
for(i in 1:length(lambda1)){

  try (
    rho[i] <- integrate(function(x) {lambda1[i]*exp(-lambda[i]*x)}, 
                        lower = 0, upper = Inf)$value)
}
rho <- rho[!is.na(rho)]
median(rho)
quantile(rho, probs = c(.025, .975))
#mean(lambda1)
plot(density(rho))
plot(log(rho), type = "l")
#max(rho)


```

